---
title: "Wordle First Letter Frequency Heat Map on a Keyboard"
subtitle: "Coding out loud, Ep 4"
author: "Sara Oberle (following Dr. Mine Çetinkaya-Rundel & Martha Aboagye)"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

```{r}
#| label: load-packages
library(tidyverse)
library(ggkeyboard) # https://github.com/sharlagelfand/ggkeyboard
```

```{r}
#| label: read-csv
wordle <- read_csv("https://raw.githubusercontent.com/tabatkins/wordle-list/main/words", col_names = "word")
```

## Inspiration

[English keyboard heatmap](https://twitter.com/monachalabi/status/772102591694209024) by [Mona Chalabi](https://twitter.com/MonaChalabi)

## Task 1: Find the first letter of Wordle words

```{r}
#| label: first-letter
wordle <- wordle |>
  mutate(first_letter = str_sub(word, start = 1, end = 1))
```

```{r}
#| label: count-first-letter
wordle_counts <- wordle |>
  count(first_letter, sort = TRUE)
```

From <https://github.com/tabatkins/wordle-list>.

## Task 2: Draw a keyboard

```{r}
ggkeyboard()
```

```{r}
#| label: ggkeyboard 
ggkeyboard(tkl, layout = "iso")
```

## Task 3: Color keyboard keys based on frequency of first letter

### Color options
dark blue - "#2D669B" - 0
light blue - "#359CA8" - 20
green - "#B8C222" - 189
yellow - "#EEDD49" - 325
orange - "#F2BC4A" - 595
dark orange - "#FB704A" - > 736
red - "#BE372C" - > 1000

### Color scale
```{r}
#| label: summary-stats
wordle_counts |>
  summarize(
    min = min(n),
    max = max(n),
    mean = mean(n),
    sd = sd(n), 
    q20 = quantile(n, 0.20), 
    q40 = quantile(n, 0.40),
    q60 = quantile(n, 0.60),
    q80 = quantile(n, 0.80)
  ) 
```

```{r}
#| label: wordle-counts 
wordle_counts |>
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 100)
```

```{r}
#| label: geom-col
wordle_counts |>
  ggplot(aes(x = first_letter , y = n)) +
  geom_col() +
  geom_hline(yintercept = 80)
```

dark blue - "#2D669B" - 0
light blue - "#359CA8" - 20
green - "#B8C222" - 189
yellow - "#EEDD49" - 325
orange - "#F2BC4A" - 595
dark orange - "#FB704A" - > 736
red - "#BE372C" - > 1000

```{r}
#| label: case-when-colors
wordle_counts <- wordle_counts |>
  mutate(
    color = case_when(
    n == 0                  ~ "#2D669B",
    n <= 20                 ~ "#359CA8", 
    n > 20 & n <= 189       ~ "#B8C222", 
    n > 189 & n <= 325      ~ "#EEDD49", 
    n > 325 & n <= 595      ~ "#F2BC4A", 
    n > 595 & n <= 1000     ~ "#FB704A", 
    n > 1000                ~ "#BE372C"
  ), 
  first_letter = str_to_upper(first_letter)
  )
wordle_counts
```


```{r}
#| label: test1
ggkeyboard(sixty_percent, palette = keyboard_palette("magic")) |>
  highlight_keys(
    keys = c("Z", "X", "C", "V"), 
    fill = c("red", "pink", "orange", "blue"), 
    alpha = 0.7, 
    colour = NA
    )
```

```{r}
#| label: test2 
ggkeyboard(sixty_percent, palette = keyboard_palette("magic")) |>
  highlight_keys(
    keys = c("C", "N", "S"), 
    fill = c("red", "pink", "orange"), 
    alpha = 0.7, 
    colour = NA
    )
```


```{r} 
#| label: read-keys-csv
keys <- read_csv("https://raw.githubusercontent.com/sharlagelfand/ggkeyboard/061f0d9084d76d576200e81439311983e1d50d93/data-raw/sixty_percent.csv")
```


```{r}
#| label: wordle-counts-ordered
wordle_counts_ordered <- keys %>% 
  filter(key_label %in% LETTERS) %>%
  select(key) %>%
  left_join(wordle_counts, by = c("key" = "first_letter"))
```

```{r}
#| label: keyboard-output
ggkeyboard(sixty_percent, palette = keyboard_palette("magic")) %>%
  highlight_keys(
    keys = wordle_counts_ordered$key, 
    fill = wordle_counts_ordered$color, 
    alpha = 0.7, 
    colour = NA
    )